# Grade Transition Patterns

Understanding how students progress through grades—whether they repeat, skip, or re-enter after time away—offers important context for interpreting cohort-based trends. While most students follow a traditional linear path, a small but meaningful share experience deviations in grade progression. These transitions can influence both cohort persistence and achievement results in later sections of the report.

This section summarizes the types of grade transitions observed among students in the current dataset.

```{r summarize_transitions}
# Generate grade transitions
transition_summary <- summarize_grade_changes(df)
transition_by_year <- transition_summary$Summary_Tables$By_Year_And_Grade
transition_by_type <- transition_summary$Summary_Tables$By_Year_And_Enrollment

```

## Summary by Academic Year

The table below shows how many students experienced each type of transition in each academic year.

```{r grade_change_table}
kable(transition_by_year, caption = "Number of Students by Transition Type and Year")

```

## Overall Transition Frequencies

This table summarizes the overall frequency of each grade transition type.

```{r grade_transition_table}
kable(transition_by_type, caption = "Overall Frequency of Grade Transition Types")

```

::: {.callout-tip title="Interpretation Tip"}

Repeaters may contribute to observed declines in cohort persistence rates, while skippers may increase apparent achievement levels for some grade-years.  
Understanding these patterns can help contextualize shifts in proficiency and mobility trends explored later in this report.

:::

## Grade Entry and Achievement Patterns

This section explores how students entered Grade `r target_grade`, including whether they stayed on track, repeated, skipped into the grade, or dropped back. It also compares achievement across these groups.

```{r analyze_transitions}
# Run the function (assumes `grade_transitions` is from summarize_grade_changes())
results <- analyze_grade_changes(transition_summary, 
                                 grade = target_grade, 
                                 achievement_var = achieve_var, 
                                 n_proficiencies = n_proficiency_levels
                                 )

```

```{r results_plots}
results$Transition_Plot

results$Achievement_Plot

kable(results$Target_Grade_Transitions_By_Year, caption = "Entry Type Counts by Academic Year")

kable(results$Skipped_Over_Grade_By_Year, caption = "Students Who Skipped Over the Grade")

kable(results$Achievement_By_Transition_And_Year, caption = "Achievement by Entry Type and Year") |>
  kable_styling(full_width = FALSE)

```

```{r summarize_grade_changes_text}
#| results: asis

# Prepare source tables
entry_tbl <- results$Target_Grade_Transitions_By_Year
ach_tbl   <- results$Achievement_By_Transition_And_Year

# Extract rows for current and previous year
entry_curr <- entry_tbl[entry_tbl$Academic_Year == to_academic_year(current_year), ]
entry_prev <- entry_tbl[entry_tbl$Academic_Year == to_academic_year(previous_year), ]
ach_curr   <- ach_tbl[ach_tbl$Academic_Year == to_academic_year(current_year), ]
ach_prev   <- ach_tbl[ach_tbl$Academic_Year == to_academic_year(previous_year), ]

# Generate entry type change summary
entry_types <- intersect(names(entry_prev), names(entry_curr))
entry_diffs <- sapply(entry_types[entry_types != "Academic_Year" & entry_types != "Total"], function(type) {
  delta <- entry_curr[[type]] - entry_prev[[type]]
  sign <- ifelse(delta > 0, "increased", ifelse(delta < 0, "decreased", "remained stable"))
  paste0(type, " ", sign, ifelse(sign == "remained stable", "", paste0(" by ", abs(delta))))
})
entry_text <- paste(entry_diffs, collapse = "; ")

# Generate achievement change summary
ach_types <- intersect(ach_curr$Entry_Type, ach_prev$Entry_Type)
ach_diffs <- sapply(ach_types, function(type) {
  p1 <- ach_prev$Percent_Proficient[ach_prev$Entry_Type == type]
  p2 <- ach_curr$Percent_Proficient[ach_curr$Entry_Type == type]
  delta <- round(p2 - p1, 1)
  sign <- ifelse(delta > 0, "increased", ifelse(delta < 0, "decreased", "remained stable"))
  paste0(type, " proficiency ", sign, ifelse(sign == "remained stable", "", paste0(" by ", abs(delta), "%")))
})
ach_text <- paste(ach_diffs, collapse = "; ")

# Combine summary
change_summary <- paste0(
  "From ", to_academic_year(previous_year), " to ", to_academic_year(current_year), 
  ", the composition of students entering Grade ", target_grade, " changed as follows: ",
  entry_text, ". In terms of achievement, proficiency rates ", ach_text, "."
)

cat(change_summary)

```
