## Grade Transition Patterns

Understanding how students progress through grades—whether they repeat, skip, or re-enter after time away—offers important context for interpreting cohort-based trends. While most students follow a traditional linear path, a small but meaningful share experience deviations in grade progression. These transitions can influence both cohort persistence and achievement results in later sections of the report.

This section summarizes the types of grade transitions observed among students in the current dataset.

```{r summarize_transitions}
# Generate grade transition summary using trajectory data
# Assumes this object is already created in 02_cohort-overview.qmd
transition_summary <- summarize_grade_changes(cohort_results$Trajectories)
transition_by_year <- transition_summary$Transition_Summary_By_Year
transition_by_type <- transition_summary$Transition_Summary_By_Type

```

## Summary by Academic Year

The table below shows how many students experienced each type of transition in each academic year.

```{r grade_change_table}
kable(transition_by_year, caption = "Number of Students by Transition Type and Year")

```

## Overall Transition Frequencies

This table summarizes the overall frequency of each grade transition type.

```{r grade_transition_table}
kable(transition_by_type, caption = "Overall Frequency of Grade Transition Types")

```

> [!TIP]
> **Interpretation Tip:**  
> Repeaters may contribute to observed declines in cohort persistence rates, while skippers may increase apparent achievement levels for some grade-years.  
> Understanding these patterns can help contextualize shifts in proficiency and mobility trends explored later in this report.


