# Grade Transition Patterns

Understanding how students progress through grades—whether they repeat, skip, or re-enter after time away—offers important context for interpreting cohort-based trends. While most students follow a traditional linear path, a small but meaningful share experience deviations in grade progression. These transitions can influence both cohort persistence and achievement results in later sections of the report.

This section summarizes the types of grade transitions observed among students in the current dataset.

```{r summarize_transitions}
# Generate grade transitions
transition_summary <- summarize_grade_changes(df)
transition_by_year <- transition_summary$Summary_Tables$By_Year_And_Grade
transition_by_type <- transition_summary$Summary_Tables$By_Year_And_Enrollment

```

## Summary by Academic Year

The table below shows how many students experienced each type of transition in each academic year.

```{r grade_change_table}
kable(transition_by_year, caption = "Number of Students by Transition Type and Year")

```

::: {.callout-tip title="Interpretation Tip"}

Repeaters may contribute to observed declines in cohort persistence rates, while skippers may increase apparent achievement levels for some grade-years.  
Understanding these patterns can help contextualize shifts in proficiency and mobility trends explored later in this report.

:::

## Overall Transition Frequencies

This table summarizes the overall frequency of each grade transition type. 

```{r grade_transition_table}
kable(transition_by_type, caption = "Overall Frequency of Grade Transition Types")

```



## Grade Entry and Achievement Patterns

This section explores how students entered Grade **`r target_grade`** by examining how the composition of the grade changed across academic years. Students are grouped based on their enrollment patterns. Some repeated the grade, meaning they were in the same grade both last year and this year. Others stayed on track by advancing one grade level as expected. A few skipped into the grade from two or more levels below, while others skipped over the grade entirely, moving from a lower grade last year to a higher grade this year. Finally, some students joined the grade without being observed in the previous year — often due to transferring in or newly entering the data system. These patterns provide insight into how student movement may influence achievement outcomes.

```{r analyze_transitions}
# Run the function
results <- analyze_grade_changes(
  dataset = df,
  grade = target_grade
)

```


```{r summarize_grade_changes_table}
# Get the summary table
entry_tbl <- results$Tables$All

kable(
  entry_tbl,
  caption = paste0("Summary of Grade ", target_grade, " Entry Types Across Academic Years"),
  col.names = gsub("_", " ", names(entry_tbl))
)

```


```{r summarize_grade_changes_text}
#| results: asis

# Extract current and previous years
year_prev <- to_academic_year(previous_year)
year_curr <- to_academic_year(current_year)

entry_tbl <- results$Tables$All
entry_prev <- entry_tbl[entry_tbl$Academic_Year == year_prev, ]
entry_curr <- entry_tbl[entry_tbl$Academic_Year == year_curr, ]
total_curr <- entry_curr$Total

# Summarize changes with labels and percents
types <- setdiff(names(entry_tbl), c("Academic_Year", "Total"))
entry_diffs <- sapply(types, function(type) {
  label <- gsub("_", " ", type)  # Replace underscores with spaces for reporting
  n_prev <- entry_prev[[type]]
  n_curr <- entry_curr[[type]]
  delta <- n_curr - n_prev
  pct <- if (!is.na(total_curr) && total_curr > 0) round(100 * abs(delta) / total_curr, 1) else NA
  direction <- if (delta > 0) "increased" else if (delta < 0) "decreased" else "remained stable"

  if (direction == "remained stable") {
    paste0("**", label, "** remained stable")
  } else {
    paste0("**", label, "** **", direction, "** by **", abs(delta), "** students",
           if (!is.na(pct)) paste0(" (_", pct, "% of cohort_)") else "")
  }
})

# Combine and print summary text
summary_text <- paste0(
  "From **", year_prev, "** to **", year_curr, "**, the composition of students entering Grade **", 
  target_grade, "** changed as follows: ", 
  paste(entry_diffs, collapse = "; "), "."
)

cat(summary_text)
```
