# Mobility Analysis

Student mobility can significantly affect cohort performance trends. This section examines how students who stayed, joined, or left the tested cohort performed on the most recent ELA assessment. In this report, **"Stay"** students were enrolled in both last year and this year (`r definitions$mobility$items$Stay`), **"Join"** students were enrolled this year but not last year (`r definitions$mobility$items$Join`), and **"Leave"** students were enrolled last year but not in the current year (`r definitions$mobility$items$Leave`). These classifications help us understand how student movement may relate to achievement outcomes.

## Summary of Key Findings

```{r compare_mobility}
# Run analysis
mobility_results <- compare_achievement_mobility(
  dataset = df, 
  current_year = current_year, 
  current_grade = target_grade,
  achievement_variable = achieve_var
)

```

```{r mobility_text}
#| results: asis

# Print the mobility text
mobility_text <- report_achievement_mobility(mobility_results)
cat(mobility_text$Paragraph_Summary)

```

## Proficiency Rates by Mobility Group

```{r comparison_plot}
# Plot the comparison plot
mobility_results$Comparison_Plot

```

```{r comparison_text}
#| results: asis

# Print an interpretation of the results
cat(mobility_text$Proficiency_By_Group)

```

## Distribution of Achievement Levels

```{r distribution_plot}
# New grouped bar plot showing achievement level breakdown
mobility_results$Achievement_Distribution_Plot

```

```{r distribution_context}
#| results: asis

# Optional: Add interpretive text if available
cat("This grouped bar chart shows how achievement levels are distributed across four categories: students who stayed or left in the previous year, and those who stayed or joined in the current year. This format helps identify shifts in proficiency related to student mobility transitions.")

```

## Achievement Level Changes for "Stay" Students

```{r stay_change_plot}
# Print the plot
mobility_results$Stay_Change_Plot

```

```{r stay_change_text}
#| results: asis

# Print the text associated with the plot
cat(mobility_text$Stay_Cohort_Change)

```

## Most Common Achievement Levels

```{r most_common}
#| results: asis

# Print the text associated with the Most Common Levels
cat(mobility_text$Most_Common_Levels)

```

