# Achievement Trends

This section examines how student achievement in **`r content_area`** has changed over time. By tracking the percentage of students who reached proficiency—based on state-defined achievement levels—we can identify long-term patterns in performance and assess progress across years.

In this analysis, we focus on **`r if (!is.null(target_grade)) paste("Grade", target_grade) else paste("Grades", paste(range(grades), collapse = "–"))`**, examining achievement trends across multiple academic years.

Proficiency is defined as scoring at or above the following achievement level(s): **`r paste(proficiency_levels, collapse = ", ")`**. In some cases, we also highlight students performing at advanced levels, which we define as scoring at or above **`r paste(advanced_levels, collapse = ", ")`**. This distinction helps us monitor not only whether students are meeting expectations but also whether high-performing students are continuing to excel.




```{r }


# Subset used in paragraph
group_summary <- df[df$GRADE == target_grade & df$YEAR_NUM == current_year, ]

# Subset used in cohort plot
group_plot <- df[df$GRADE == target_grade & df$YEAR == paste0(current_year - 1, "_", current_year), ]

# Compare sample sizes and counts
cat("Summary group size: ", nrow(group_summary), "\n")
cat("Cohort plot group size: ", nrow(group_plot), "\n")

# Table of proficiency levels
cat("\nProficiency table (summary):\n")
print(table(group_summary$PROFICIENCY_LEVELS))

cat("\nProficiency table (plot):\n")
print(table(group_plot$PROFICIENCY_LEVELS))

```



## Proficiency Trends by Synthetic Cohorts


```{r cohort_proficiency}

# Plot the synthetic cohort trends
cohort_plot <- plot_cohort2_proficiency(
  dataset = df, 
  year_range = available_years, 
  grade_range = grade_range,
  n_proficiencies = n_proficiency_levels, 
  achievement = achieve_var
)

cohort_plot$Plot
cohort_plot$Table

```