# Appendices

## Appendix: Cohort-Based Achievement Trends (Visual)

The chart below shows how proficiency rates have changed over time for each tested grade level. By aligning results from the same grade across multiple years, this visualization highlights patterns in student performance—such as overall improvement, decline, or stability at each grade.

This view helps contextualize performance across the system and may reveal whether certain grades consistently show higher or lower proficiency levels compared to others.

```{r proficiency_plot}
# Reuse previously created plot object
cohort_plot$Plot

```

```{r advanced_plot}
# Reuse previously created plot object
advanced_plot$Plot

```

## Appendix: Individual Cohort Heatmaps

```{r cohort_heatmaps, fig.align='center', fig.width=8, fig.height=6}
#| results: asis

# Loop through each heatmap in the list and print it with a header
for (name in names(cohort_results$Heatmaps)) {
  if (name == "Combined") next  # Skip the combined heatmap
  
  cat("### Cohort Join Year: ", name, "\n\n")
  print(cohort_results$Heatmaps[[name]])
  cat("\n\n")
}

```


