---
title: "State Level Mobility Report"
author: "Brian Harrold"
format: 
  html: 
    self-contained: true
execute: 
  cache: false
---

## Overview

This short report provides a snapshot of student mobility trends in the state based on recent available data. The mobility of concern is for those students who enter or exit the state based on the Grade 3 assessment from 2019-2020 and Grade 4 assessment from 2020-2021.

## Key Findings

The alluvial diagram below shows the student mobility of 3rd graders from 2019-2020 to 4th graders from 2020-2021.

```{r alluvial}
#| echo: FALSE

library(cohortED)
library(ggplot2)
library(cowplot)

grade3 = plot_alluvial_mobility(dataset = math, start_year = "2019_2020", start_grade = 3, 
                       print_table = FALSE, data_out = TRUE)

```

Student mobility in 3rd grade appears to be more highly associated with the non-white population.

```{r ethnicity_table}
#| echo: FALSE

mobility_ethnicity = table(grade3$MOBILITY_STATUS, grade3$ETHNICITY)

knitr::kable(round(prop.table(mobility_ethnicity, margin = 2), 3))

```


## Impact of Student Mobility


Student mobility tends impact performance on assessments. The both the students who left and the students who joined had a higher proportion of students who were not proficient. Additionally, the students a higher proportion of the students who left had unsatisfactory performance on the Grade 3 assessment. While there were a similar proportion of students on the Grade 4 assessment in both the Stay and Join groups, the students who joined tended to not have the same proportion of advanced scores. The performance of students who stayed seemed to be very similar between Grade 3 and Grade 4.


```{r student_performance}
#| echo: FALSE

grade3.x = grade3[!is.na(grade3$ACHIEVEMENT_LEVEL.x) & grade3$ACHIEVEMENT_LEVEL.x != "No Score", ]
grade3.y = grade3[!is.na(grade3$ACHIEVEMENT_LEVEL.y) & grade3$ACHIEVEMENT_LEVEL.y != "No Score", ]

table1 = table(grade3.x$MOBILITY_STATUS, grade3.x$ACHIEVEMENT_LEVEL.x)
df1 = data.frame(round(prop.table(table1, margin = 1), 3))
df1 = df1[!is.nan(df1$Freq), ]
df1$Var2 = factor(df1$Var2, 
                  levels = c("Advanced", "Proficient", "Partially Proficient", "Unsatisfactory"))

table2 = table(grade3.y$MOBILITY_STATUS, grade3.y$ACHIEVEMENT_LEVEL.y)
df2 = data.frame(round(prop.table(table2, margin = 1), 3))
df2 = df2[!is.nan(df2$Freq), ]
df2$Var2 = factor(df2$Var2, 
                  levels = c("Advanced", "Proficient", "Partially Proficient", "Unsatisfactory"))

plot1 = ggplot(df1, aes(x = Var1, y = Freq, fill = Var2)) + 
  geom_bar(stat = "identity") + 
  labs(title = "2019-2020", y = "Value", x = "Mobility Status", fill = "Achievement Level") + 
  theme_minimal() +
  scale_fill_manual(values = c(
    "Advanced" = "#2E8B57",
    "Proficient" = "#66CDAA",
    "Partially Proficient" = "#E57373",
    "Unsatisfactory" = "#B22222"
  ))

plot2 = ggplot(df2, aes(x = Var1, y = Freq, fill = Var2)) + 
  geom_bar(stat = "identity") + 
  labs(title = "2020-2021", y = "Value", x = "Mobility Status", fill = "Achievement Level") + 
  theme_minimal() +
  scale_fill_manual(values = c(
    "Advanced" = "#2E8B57",
    "Proficient" = "#66CDAA",
    "Partially Proficient" = "#E57373",
    "Unsatisfactory" = "#B22222"
  ))

plot_grid(plot1, plot2, ncol = 2)

```



## Data Notes

This is an example report based on the sgpData_Long that is contained in the SGPData package, which has been subset to only include the data associated the **Mathematics** Content Area.





