---
title: "Definitions"
format:
  html:
    toc: true
    toc-depth: 2
---

## Glossary of Key Terms

This page provides plain-language definitions for key concepts used across the `cohortED` reporting package. These definitions support consistency, clarity, and accessibility in automated report narratives.

You can include any of these definitions in your own custom reports — each one is stored as part of the `definitions` object in `cohortED`, and is ready to be reused anywhere Quarto or R Markdown text is rendered.

While this page shows the `bullet_text` version (pre-formatted for markdown), each definition list also includes a structured `items` field containing named key-value pairs for programmatic access.

---

## Definitions

```{r setup, include=FALSE}
library(cohortED)
data("definitions")

```

<details>
<summary style="font-size: 1.2em; font-weight: bold;">Mobility</summary>



```{r echo=FALSE, results='asis'}
cat(definitions$mobility$bullet_text)

```


</details> <details> <summary style="font-size: 1.2em; font-weight: bold;">Achievement Levels</summary>


```{r echo=FALSE, results='asis'}
cat(definitions$achievement_levels$bullet_text)

```



</details> <details> <summary style="font-size: 1.2em; font-weight: bold;">Trajectory</summary>


```{r echo=FALSE, results='asis'}
cat(definitions$trajectory$bullet_text)

```


</details> <details> <summary style="font-size: 1.2em; font-weight: bold;">Transitions</summary>


```{r echo=FALSE, results='asis'}
cat(definitions$transitions$bullet_text)

```



</details>